---
title: "Coursera_MachLearning"
author: "Manuel Kiewisch"
date: "Wednesday, July 22, 2015"
output: html_document
---

##Session Info (Platform/ Version Information)
```{r}
sessionInfo()
```

##Purpose  
- predict exercise data from personal activity devices
- predict if available data per exercise has been collected from a ''correct'' or ''incorrect'' performance
- report on the model building process and describe the expected out-of-sample error  

##Data  
- based on the ''Groupware'' data set available at: http://groupware.les.inf.puc-rio.br/har  

##Data Preparation  
- a training and test data set have been combiled and made available at: 
- https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv  
- https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
- the memory available to R is set to 4gb
- all variables containing missing values have been remove 
- further, all purely descriptive variables (variables 1--7, i.e. name, id, timestamps) have been removed since they are not expected to be of any predictive value
- the predictor variable $classe is transformed to a factor variable
- the remaining variables are scanned for near-Zero-Variables and near-Zero-Variables are removed  

```{r, echo=FALSE}
memory.limit(size=4000)
library(caret)
library(ggplot2)
setwd("C:\\Users\\ElitebookHP\\Desktop\\coursera_machLearning")
training1 <- read.table("pml-training.csv",as.is = TRUE, na.strings=c("NA", "-", "?","","#DIV/0!"),sep = ",",header = TRUE)
training <- training1[,colSums(is.na(training1))==0]
training <- training[,c(-1,-2,-3,-4,-5,-6,-7)]
training$classe <- as.factor(training$classe)
nsv <- nearZeroVar(training,saveMetrics=TRUE)
training <- training[,c(row.names(nsv[nsv$nzv==FALSE,]))]
testing1 <- read.table("pml-testing.csv",as.is = TRUE, na.strings=c("NA", "-", "?","","#DIV/0!"),sep = ",",header = TRUE)
testing <- testing1[,colSums(is.na(testing1))==0]
testing <- testing[,c(-1,-2,-3,-4,-5,-6,-7)]
```
  
##Data Splicing  
- the original training set sampled to 9000 cases in order to overcome hardware related memory issues
- the sampled training set is then partitioned into 60/40 into a training and test set to appraise the out-of-sample-error  
```{r, echo=FALSE}
training <- [sample(nrow(training), 9000), ]
inTrain <- createDataPartition(y=training$classe, p=0.3,list=FALSE)
training <- training[inTrain,]
test <- training[-inTrain,]
```

## Model Building  
- a 'Random Forest' modeling procedure is used to create the prediction model  
- cross-validation is used to select the model with highest accuracy  
- pre-processing is used to adjust the remaining variables, including a Principal-Component-Analysis to identify valuable predictor groups
- the following graph displays the result of the principal component analysis for components 1 and 2 for the training data partition  
```{r, echo=FALSE}
modelfit <- train(classe ~.,method="rf",data=training,trControl=trainControl(method="cv"),preProcess=c("center","scale","pca"),prox=TRUE)
pca <- prcomp(training[, -c(53)], scale.=TRUE)
biplot(pca, scale=0, cex=.6)
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
```

## Model Results and Out-of-sample error  
- the following table and plot display the model results and accurracy including the out-of-sample error (inverse accurracy)  
```{r, echo=FALSE}
print(modelfit)
predictions <- predict(modelfit,newdata=test)
confusionMatrix(predictions,test$classe)

pred <- data.frame(predictions,test$classe)
p <- ggplot(pred, aes(x = predictions, y = test$classe))
p <- p + geom_jitter(position = position_jitter(width = 0.25, height = 0.25))
p
```

## Predictions  
- in the following, the model is used to predict 20 values from a testing set  
- the prediction results are saved in 20 separate text files  
```{r, echo=FALSE}
answers <- predict(modelfit,newdata=testing)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)
```
